### 2018-12-22 Sections/index.tex
indexing
more indexing
Indexed QM
quote

@@

\index{Random graph \seeonly{Percolation graph}|gobbleone}

@@

\index{Redundancy code \seeonly{3-qubit code}|gobbleone}

@@

\index{Phase-flip code \seeonly{3-qubit code}|gobbleone}

@@

@@

\index{Period-finding  @\also{Shor's algorithm}|gobbleone}
\index{Robustness @\also{Network robustness}|gobbleone}
\index{Message digests @\also{Checksum}|gobbleone}
\index{Collisions @\also{Hash collisions}|gobbleone}
\index{Checksum @\also{Hash functions}|gobbleone}
\index{Black box @\also{Oracle}|gobbleone}
\index{Thermal noise @\also{Errors}|gobbleone}
\index{Synthetic aperture @\also{Quantum-enhanced telescopy}|gobbleone}
\index{Ledger @\also{Distributed ledger}|gobbleone}
\index{Integer factorisation @\also{Shor's algorithm}|gobbleone}
\index{Human genome project @\also{Genome}|gobbleone}
\index{Enigma machines @\also{Quantum Enigma machines}|gobbleone}
\index{Efficiency @\also{Loss channel, detector efficiency, computational efficiency}|gobbleone}
\index{Cryptanalysis @\also{Differential cryptanalysis, linear cryptanalysis, frequency analysis}|gobbleone}
\index{Quantum Moore's Law @\also{Moore's Law}|gobbleone}
\index{Toric code @\also{Topological codes, surface codes}|gobbleone}
\index{Surface codes @\also{Topological codes}|gobbleone}
\index{Distributed unitary error averaging @\also{Unitary error averaging}|gobbleone}
\index{Nano-satellites @\also{Micro-satellites}|gobbleone}
\index{Micro-satellites @\also{Nano-satellites}|gobbleone}
\index{Two-channel two-port switches @\also{Optical routers}|gobbleone}
\index{Multiplexers @\also{Optical routers}|gobbleone}
\index{Demultiplexers @\also{Optical routers}|gobbleone}
\index{Linear multiplexers \& demultiplexers @\also{Optical routers}|gobbleone}
\index{Pyramid multiplexers \& demultiplexers @\also{Optical routers}|gobbleone}
\index{Single-channel multi-port switches @\also{Optical routers}|gobbleone}
\index{Multi-channel multi-port switches @\also{Optical routers}|gobbleone}
\index{Quantum error correction (QEC) @\also{3-qubit code, 5-qubit code, 7-qubit code, 9-qubit code, topological codes, unitary error averaging}|gobbleone}
\index{Homomorphic encryption @\also{Encrypted quantum computation}|gobbleone}
\index{Blind quantum computation @\also{Encrypted quantum computation}|gobbleone}
\index{Phase-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Polarisation-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Displacement-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Non-optical encodings @\also{Matter qubits}|gobbleone}
\index{Time-dependent quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Distance to horizon @\also{Line-of-sight}|gobbleone}
\index{Line-of-sight @\also{Distance to horizon}|gobbleone}
\index{Micro-cluster states @\also{Star micro-cluster states, plus micro-cluster states, linear micro-cluster states, snowflake micro-cluster states}|gobbleone}
\index{Star micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Plus micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Linear micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Snowflake micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Single-qubit quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Time-dependent quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Quality of service (QoS) @\also{Quantum error correction, entanglement purification}|gobbleone}
\index{Anonymous surveys @\also{Quantum voting}|gobbleone}
\index{Binary voting @\also{Quantum voting}|gobbleone}

>>>>

\index{Kronecker product \seeonly{Tensor product}|gobbleone}

@@

\index{Random!Graph \seeonly{Percolation graph}|gobbleone}

@@

\index{Redundancy!Code \seeonly{3-qubit code}|gobbleone}

@@

\index{Phase!Flip code \seeonly{3-qubit code}|gobbleone}

@@

\index{Universal Turing machines\seeonly{Turing machines}|gobbleone}

@@

\index{Period-finding  @\also{Shor's algorithm}|gobbleone}
\index{Robustness @\also{Network robustness}|gobbleone}
\index{Message digests @\also{Checksum}|gobbleone}
\index{Collisions @\also{Hash collisions}|gobbleone}
\index{Checksum @\also{Hash functions}|gobbleone}
\index{Black box @\also{Oracle}|gobbleone}
\index{Thermal noise @\also{Errors}|gobbleone}
\index{Synthetic aperture @\also{Quantum-enhanced telescopy}|gobbleone}
\index{Ledger @\also{Distributed ledger}|gobbleone}
\index{Integer factorisation @\also{Shor's algorithm}|gobbleone}
\index{Human genome project @\also{Genome}|gobbleone}
\index{Enigma machines @\also{Quantum Enigma machines}|gobbleone}
\index{Efficiency @\also{Loss channel, detector efficiency, computational efficiency}|gobbleone}
\index{Cryptanalysis @\also{Differential cryptanalysis, linear cryptanalysis, frequency analysis}|gobbleone}
\index{Quantum Moore's Law @\also{Moore's Law}|gobbleone}
\index{Toric code @\also{Topological codes, surface codes}|gobbleone}
\index{Surface codes @\also{Topological codes}|gobbleone}
\index{Distributed unitary error averaging @\also{Unitary error averaging}|gobbleone}
\index{Nano-satellites @\also{Micro-satellites}|gobbleone}
\index{Micro-satellites @\also{Nano-satellites}|gobbleone}
\index{Two-channel two-port switches @\also{Optical routers}|gobbleone}
\index{Multiplexers @\also{Optical routers}|gobbleone}
\index{Demultiplexers @\also{Optical routers}|gobbleone}
\index{Linear multiplexers \& demultiplexers @\also{Optical routers}|gobbleone}
\index{Pyramid multiplexers \& demultiplexers @\also{Optical routers}|gobbleone}
\index{Single-channel multi-port switches @\also{Optical routers}|gobbleone}
\index{Multi-channel multi-port switches @\also{Optical routers}|gobbleone}
\index{Quantum error correction (QEC) @\also{3-qubit code, 5-qubit code, 7-qubit code, 9-qubit code, topological codes, unitary error averaging}|gobbleone}
\index{Homomorphic encryption @\also{Encrypted quantum computation}|gobbleone}
\index{Blind quantum computation @\also{Encrypted quantum computation}|gobbleone}
\index{Phase!Key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Polarisation-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Displacement-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Non-optical encodings @\also{Matter qubits}|gobbleone}
\index{Time-dependent quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Distance to horizon @\also{Line-of-sight}|gobbleone}
\index{Line-of-sight @\also{Distance to horizon}|gobbleone}
\index{Micro-cluster states @\also{Star micro-cluster states, plus micro-cluster states, linear micro-cluster states, snowflake micro-cluster states}|gobbleone}
\index{Star micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Plus micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Linear micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Snowflake micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Single-qubit quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Time-dependent quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Quality of service (QoS) @\also{Quantum error correction, entanglement purification}|gobbleone}
\index{Anonymous surveys @\also{Quantum voting}|gobbleone}
\index{Binary voting @\also{Quantum voting}|gobbleone}

### 2018-12-22 Sections/intro_to_QM.tex
quote

\dropcap{T}{o} make the upcoming discourse accessible to those from a non-quantum background, we will briefly level the playing-field by presenting an overview of the mathematical formulation of quantum mechanics, all of which fits within the confines of familiar basic linear algebra\index{Linear algebra}\index{Quantum mechanics}.

@@

The laws of quantum mechanics are inherently different to those of classical mechanics, and bring with them their own unique notation for:
\begin{itemize}
	\item The representation of quantum states.
	\item The evolution of quantum systems.
	\item The measurement of quantum states.
\end{itemize}
We review all of these formalisms in the upcoming sections, and present the necessary mathematical framework to enable following the derivations presented here.

@@

\dropcap{C}{lassical} and quantum physics are fundamentally different in a number of extremely important ways. These differences are what allow quantum mechanics to enable the entire new field of quantum technologies, most notably quantum computing, which offers massive computational gains over traditional classical computing. The same applies to the novel and groundbreaking promises of the other key quantum technologies. All of these gains are achieved by exploiting these unique new characteristics offered by quantum systems, that are physically impossible in a classical context.

@@

All of these phenomena will become more clear and intuitive as we present the relevant mathematical constructions in the upcoming sections.

@@

\dropcap{A}{} quantum state is a mathematical description of a quantum object, fully characterising its quantum mechanical properties, much like classical attributes like momentum and position characterise the state of classical objects. This captures both the wave- and particle-like properties of quantum objects, given the quantum principle of \textit{wave-particle duality}\index{Wave-particle duality}. The key mathematical difference is that quantum objects exhibit unique, non-classical properties like superposition and entanglement.

@@

In this book we will largely be dealing with qubits\index{Qubits}, the simplest unit of quantum information, which are supported by exactly two basis states, denoted $\ket{0}$ and $\ket{1}$. A qubit can therefore be expressed as,
\begin{align}
\ket\psi &= \left(\begin{matrix}{}
	\alpha \\
	\beta
\end{matrix}\right)\nonumber\\
&= \alpha\ket{0} + \beta\ket{1}.
\end{align}
These are referred to as the \textit{logical basis states}\index{Logical basis states}, which are treated abstractly, detached from a particular physical realisation (e.g by associating them with electron energy levels, photon polarisation, or indeed \textit{any} two-level quantum system), much in the same way that classical bits are treated abstractly, detached from physical reality.

@@

For two generic state vectors,
\begin{align}
	\ket\psi = \sum_n \alpha_n \ket{n},\nonumber\\
	\ket\phi = \sum_n \beta_n \ket{n},
\end{align}
the \textit{overlap}\index{Overlap} between them is defined as,
\begin{align}
\braket{\psi|\phi} = \sum_n \alpha_n\beta_n^*,
\end{align}
which is mathematically equivalent to their vector dot product\index{Dot product}. The object $\braket{\cdot|\cdot}$ is referred to as a \textit{braket}\footnote{We never quite appreciated the corny sense of humour of the author of the terms \textit{bra}, \textit{ket}, and \textit{braket}. But somehow it stuck.}\index{Braket}. Owing to the cyclic property of the trace operator\index{Trace operator}, it follows that the trace,
\begin{align}
\mathrm{tr}(\ket\psi\bra\phi)=\braket{\psi|\phi},	
\end{align}
is a state overlap. This quantity has the useful intuitive interpretation as the \textit{distinguishability}\index{Distinguishability} between quantum states, which plays an important role in measurement theory and the theory of quantum information. Thus, the normalisation condition from Eq.~(\ref{eq:state_norm_cond}) can equivalently be written,
\begin{align}\index{Normalisation}
\braket{\psi|\psi} = 1.	
\end{align}
Because basis states are orthonormal, this implies that for basis states $\ket{m}$ and $\ket{n}$ we have the trivial delta function overlap identity,
\begin{align}
	\braket{m|n}=\delta_{m,n}.
\end{align}

@@

The intuitive physical interpretation of density matrices is as follows:
\begin{itemize}
\item Rows and columns enumerate the basis states of the Hilbert space.
\item Diagonal elements represent the classical probabilities\index{Classical probabilities} of the respective basis states.
\item Off-diagonal elements represent \textit{coherences}\index{Coherences} between pairs of basis states, i.e whether they exist as a classical probability distribution (when the coherence is zero) or as a quantum superposition (when non-zero). The $\hat\rho_{m,n}$ matrix element represents the coherence between basis states $\ket{m}$ and $\ket{n}$.
\end{itemize}

@@

Applying our measurement identities, we obtain measurement probabilities,
\begin{align}
p_0 = |\braket{0|\psi}|^2 = \frac{1}{2},\nonumber\\
p_1 = |\braket{1|\psi}|^2 = \frac{1}{2},
\end{align}
and respective post-measurement projected states,
\begin{align}
\ket{\psi_0} &= \sqrt{2}\ket{0}\braket{0|\psi} = \ket{0},\nonumber\\
\ket{\psi_1} &= \sqrt{2}\ket{1}\braket{1|\psi} = \ket{1}.
\end{align}
Thus, we see that with 50:50 probability the measurement randomly collapses the measured state onto the $\ket{0}$ or $\ket{1}$ states.

>>>>

\famousquote{I think I can safely say that nobody understands quantum mechanics.}{Richard Feynman}
\\

\dropcap{T}{o} make the upcoming discourse accessible to those from a non-quantum background, we will briefly level the playing-field upfront by presenting an overview of the key concepts and mathematical formulation of quantum mechanics, all of which fits within the confines of familiar basic undergraduate linear algebra\index{Linear algebra}\index{Quantum mechanics}.

@@

However, rather than just present the mathematical formulation of quantum mechanics and expect you to just take the maths as the truth and the way, we will try as best we can to inject some intuition and meaningful interpretation of the mathematics into our tutorial, so that hopefully the reader is able to walk away with not just the ability to follow basic quantum mechanical derivations, but also grasp the intuition of the underlying physical reality.

The laws of quantum mechanics are inherently different to those of classical mechanics in virtually every way, and bring with them their own unique notation for:
\begin{itemize}
	\item The representation of quantum states.
	\item The evolution of quantum systems.
	\item The measurement of quantum states.
\end{itemize}
We review all of these formalisms in the upcoming sections, and present the necessary mathematical framework to enable following the derivations presented here.

@@

\dropcap{C}{lassical} and quantum physics are fundamentally different in a number of extremely important ways. These differences are what allow quantum mechanics to enable the entire new field of quantum technologies, most notably quantum computing, which offers massive computational gains over traditional classical computing. The same applies to the novel and groundbreaking promises of the other key quantum technologies, such as quantum cryptography, quantum teleportation, and quantum-enhanced metrology. All of these gains are achieved by exploiting these unique new characteristics offered by quantum systems, that are physically impossible in a classical context.

@@

All of these phenomena will become more clear and intuitive as we introduce the basic relevant mathematical constructions in the upcoming sections.

@@

\dropcap{A}{} quantum state is a mathematical description of a quantum object, fully characterising its quantum mechanical properties, much like classical attributes like momentum and position characterise the state of classical objects. This captures both the wave- and particle-like properties of quantum objects, given the quantum principle of \textit{wave-particle duality}\index{Wave-particle duality}. The key mathematical difference is that quantum objects exhibit unique, non-classical properties like superposition and entanglement. This requires introducing new constructs for their mathematical representation.

@@

In this book we will largely be dealing with qubits\index{Qubits}, the simplest unit of quantum information, which are supported by exactly two basis states, denoted $\ket{0}$ and $\ket{1}$. A qubit can therefore be expressed as,
\begin{align}
\ket\psi &= \left(\begin{matrix}{}
	\alpha \\
	\beta
\end{matrix}\right)\nonumber\\
&= \alpha\ket{0} + \beta\ket{1}.
\end{align}
These are referred to as the \textit{logical basis states}\index{Logical basis states}, which are treated abstractly, detached from a particular physical realisation (e.g by associating them with electron energy levels, photon polarisation, or indeed \textit{any} two-level quantum system), much in the same way that classical bits are treated abstractly as a unit of information, detached from physical reality.

@@

For two generic state vectors, supported by the same basis,
\begin{align}
	\ket\psi = \sum_n \alpha_n \ket{n},\nonumber\\
	\ket\phi = \sum_n \beta_n \ket{n},
\end{align}
the \textit{overlap}\index{Overlap} between them is defined as,
\begin{align}
\braket{\psi|\phi} = \sum_n \alpha_n\beta_n^*,
\end{align}
which is mathematically equivalent to their vector dot product\index{Dot product}. The object $\braket{\cdot|\cdot}$ is referred to as a \textit{braket}\footnote{We never quite appreciated the corny sense of humour of the author of the terms \textit{bra}, \textit{ket}, and \textit{braket}. But somehow it stuck.}\index{Braket}. Owing to the cyclic property of the trace operator\index{Trace operator}, it follows that the trace,
\begin{align}
\mathrm{tr}(\ket\psi\bra\phi)=\braket{\psi|\phi},	
\end{align}
is a state overlap. This quantity has the useful intuitive interpretation as the \textit{distinguishability}\index{Distinguishability} between quantum states, which plays an important role in measurement theory and the theory of quantum information. Thus, the normalisation condition from Eq.~(\ref{eq:state_norm_cond}) can equivalently be written,
\begin{align}\index{Normalisation}
\braket{\psi|\psi} = 1.	
\end{align}
Because basis states are orthonormal\index{Orthonormality}, this implies that for basis states $\ket{m}$ and $\ket{n}$ we have the trivial delta function overlap identity,
\begin{align}
	\braket{m|n}=\delta_{m,n}.
\end{align}

@@

The intuitive physical interpretation of density matrices is as follows:
\begin{itemize}
\item Rows and columns enumerate the basis states of the Hilbert space.
\item Diagonal elements represent the classical probabilities\index{Classical probabilities} of the respective basis states.
\item Off-diagonal elements represent \textit{coherences}\index{Coherences} between pairs of basis states, i.e whether they exist as a classical probability distribution (when the coherence is zero) or as a quantum superposition (when non-zero). The $\hat\rho_{m,n}$ matrix element represents the degree of coherence between basis states $\ket{m}$ and $\ket{n}$.
\end{itemize}

@@

Applying our measurement identities, we obtain measurement probabilities,
\begin{align}
p_0 = |\braket{0|\psi}|^2 = \frac{1}{2},\nonumber\\
p_1 = |\braket{1|\psi}|^2 = \frac{1}{2},
\end{align}
and respective post-measurement projected states,
\begin{align}
\ket{\psi_0} &= \sqrt{2}\ket{0}\braket{0|\psi} = \ket{0},\nonumber\\
\ket{\psi_1} &= \sqrt{2}\ket{1}\braket{1|\psi} = \ket{1}.
\end{align}

Thus, we see that with 50:50 probability the measurement randomly collapses the measured state onto the $\ket{0}$ or $\ket{1}$ states.
